Commutative Character Sheaves

### 2015-11-03 followup.tex
Fixing/skipping some tikzcd problems
Formating of equalizer arrows, moving Theorem 1.5 into section 1

\section{Recollections and definitions}

@@

The first step in this process is to rule out character sheaves as above.
The following proposition will be proven in Section \ref{sec:descent}.
\begin{proposition}
Suppose $G$ is a smooth group scheme and $\cs{L} \in \CS(G)$ is a linear character sheaf on $G$.
Then the restriction of $\cs{L}$ to $G\der$ is trivial if and only if $\cs{L} \cong q^*(\cs{L}\ab)$ for some
character sheaf $\cs{L}\ab$ on $G\ab$.
\end{proposition}

@@

\begin{proposition}
A linear character sheaf $(\cs{L}, \mu)$ on a smooth group scheme $G$ is commutative if and only if
\begin{enumerate}
\item $\cs{L} \cong q^* \cs{L}\ab$ for some character sheaf $\cs{L}\ab$ on $G\ab$,
\item the following diagram of Weil sheaves on $G \times G$ commutes.
  \[
  \begin{tikzcd}[row sep=30]
   m^*\cs{L} \arrow{d}[swap]{(q\times q)^*\xi\ab} \arrow{r}{\mu} & \cs{L}\boxtimes\cs{L} \arrow{d}{\vartheta}\\
   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} &  \theta^*(\cs{L}\boxtimes\cs{L})
  \end{tikzcd}
  \]
\end{enumerate}
\end{proposition}

@@

Let $G$ be as above.
Let $H$ be a group scheme and write $n : H \times H \to H$ for the multiplication morphism.
Let $a : H \times G\to G$ be a group action compatible with the group structure on $G$ and write $p : H\times G \to G$ for projection. 
Consider the morphisms
\[
\begin{tikzcd}
H \times H \times G 
\arrow[shift left=4]{r}{b_1} 
\arrow{r}{b_2} 
\arrow[shift right=4]{r}{b_3} 
& H \times G 
\arrow[shift left]{r}{a}
\arrow[shift right]{r}[swap]{p}
 & G 
\end{tikzcd}
\]
defined by $b_1(h_1,h_2,g) = (n(h_1,h_2),g)$,  $b_2(h_1,h_2,g) = (h_1,a(h_2,g))$ and $b_3(h_1,h_2,g) = (h_2,g)$.
Then, $a\circ b_1 = a\circ b_2$ and $a\circ b_3 = p\circ b_2$ and $p\circ b_1 = p\circ b_3$.
An $H$-equivariant Weil local sytem on $G$ is a Weil local system $\cs{L}$ on $G$ together with an isomorphism  
\[
\nu : a^*\cs{L} \to p^*\cs{L}
\] 
of Weil local systems on $H\times G$ such that the following diagram of isomorphisms of local systems on $H\times H\times G$ commutes,
\[
\begin{tikzcd}
\ &  \arrow{dl}[swap]{a\circ b_1 = a\circ b_2}  b_2^*  a^*\, \mathcal{L} \arrow{rr}{b_2^*(\nu)} && b_2^* p^*\, \mathcal{L} \arrow{dr}{p\circ b_2 = a\circ b_3} & \\
b_1^* a^*\, \mathcal{L} \arrow{dr}[swap]{b_1^*(\nu)} &&&&   \arrow{dl}{b_3^*(\nu)} b_3^* a^*\, \mathcal{L} \\
& b_1^* p^*\, \mathcal{L}  && \arrow{ll}[swap]{p\circ b_3 = p\circ b_1} b_3^* p^*\, \mathcal{L}, & 
\end{tikzcd}
\]
and $s^*(\nu) = \id_{\cs{L}}$.
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations above; so, for example, the top left isomorphism comes from $b_1^* a^* \iso b_2^* a^*$ since $a\circ b_1 = a\circ b_2$.
\iffalse
We remark that, on stalks, this condition is ...
\[
\begin{tikzcd}
\ &  \cs{L}_{x\cdot(y\cdot z)} \arrow{r}{\nu_{x,y\cdot z}} & \cs{L}_{y\cdot z} & \\
\cs{L}_{(x\cdot y)\cdot z} \arrow{dr}[swap]{\nu_{x\cdot y,z}} \arrow[equal]{ur} &&& \arrow[equal]{ul}  \arrow{dl}{\nu_{y,z}} \cs{L}_{y\cdot z} \\
& \cs{L}_z \arrow[equal]{r} & \cs{L}_z & 
\end{tikzcd}
\]
\fi
%
Morphisms of $H$-equivariant Weil local systems $(\cs{L},\nu)\to (\cs{L}',\nu')$ are morphisms of Weil local systems $\alpha: \cs{L}\to \cs{L}'$ for which the diagram
\[
\begin{tikzcd}
\arrow{d}[swap]{\nu} a^*\cs{L} \arrow{r}{a^*(\alpha)} & a^*\cs{L}' \arrow{d}{\nu'} \\
p^*\cs{L} \arrow{r}{p^*(\alpha)} & p^*\cs{L}'
\end{tikzcd}
\]
commutes.
This defines $\Loc_H(G)$, the category of $H$-equivariant Weil local systems on $G$.

@@

Now suppose $q : G \to G'$ is a regular epimorphism with kernel pair $(a,p)$ in the category of schemes
\[
\begin{tikzcd}
H\times G
 \arrow[shift left]{r}{a}
  \arrow[shift right,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G'
\end{tikzcd}
\]
Let $\cs{L}'$ be a linear character sheaf on $G'$.
Consider the functor
\[
q^* : \CS(G') \to \CS(G)
\]
given on objects by $(\cs{L}',\mu') \mapsto (q^*\cs{L}', (q\times q)^* \mu', \nu_{\cs{L}'})$; cf [Dictionary].
Then $q^*\cs{L}'$ is an $H$-equivariant linear character sheaf on $G$ when equipped with the canonical isomorphism
\[
\nu_{\cs{L}'} : a^* q^* \cs{L}' \mathop{\longrightarrow}\limits^{\iso} (q\circ a)^* \cs{L}' = (q\circ p)^*\cs{L}' \mathop{\longrightarrow}\limits^{\iso} p^* q^*\cs{L}.
\]
This simple observation shows that the functor $q^* : \CS(G') \to \CS(G)$  factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &&\arrow{ll}[swap]{q^*} 
\CS(G') \arrow{dl}{\text{`comparison'}}\\
& \arrow{ul}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]

@@

Suppose there is some $f : H\times G\to G\times G$ such that the following diagram commutes.
\[
\begin{tikzcd}
&  \arrow{dl}[swap]{a} H \times G \arrow{dr}{p} \arrow{dd}[swap]{f} & \\
G & & G \\
& \arrow{ul}{m} G \times G \arrow{ur}[swap]{p} & 
\end{tikzcd}
\]
Then $f(h,g) = (f_1(h,g),p)$ for $f_1 : H\times G \to G$.
Let $\CS'(G)$ be the full subcategory of $\CS(G)$ consisting of those $(\cs{L},\mu,\nu) \in \CS_H(G)$ such that $f_1^*\cs{L} \iso (\EE)_{H\times G}$.

@@

\subsection{Proof of Proposition~1.2}

We now consider a special case of the lemmas above. 
Suppose $H = G\der$ and the action $a$ is the restriction of $m$ to $G\der\times G$.
Then $f : G\der \times G\to G\times G$ is simply the inclusion and $f_1: G\der\times G \to G$ is projection to the first coordinate followed by inclusion $G\der \hookrightarrow G$.

As we have seen above, the fppf quotient $q : G\to G\ab = G/G\der$ exists in schemes and is a $G\der$-torsor in the fppf topology. 
%(In fact, it's a trivial torsor in that topology.)
It follows from Lemma~[] that the comparison functor $\CS(G\ab)\to \CS_{G\der}(G)$ is an equivalence.
On the other hand, by Lemma~[], the forgetful functor $\CS_{G\der}(G)\to \CS(G)$ defines an equivalence $\CS_{G\der}(G)\to \CS'(G)$, where $\CS'(G)$ is the category of $(\cs{L},\mu,\nu)\in \CS_{G\der}(G)$ such that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$.
It now follows that $q^* : \CS(G\ab) \to \CS(G)$ is fll and faithful with essential image $\CS'(G)$.
This concludes the proof of Proposition~1.2.

\subsection{Commutative character sheaves}

Move this statement to Section 1?

\begin{proof}
Combining Lemmas~[] and [] we see that $q^* : \CS(G\ab) \to \CS(G)$ is full and faithful.
% with essential image $\CS'(G)$, the category of linear character sheaves $\cs{L}$ on $G$ for which $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$. 
By Definition~[], this induces an equivalence between $\CCS(G\ab)$ and $\CCS(G)$. 
\end{proof}

\subsection{Proof of Proposition~1.4}

>>>>

\section{Recollections and definitions} \label{sec:defs}

@@

The first step in this process is to rule out character sheaves as above.
The following proposition will be proven in Section \ref{sec:descent}.
\begin{proposition} \label{prop:Gder_triv}
Suppose $G$ is a smooth group scheme and $\cs{L} \in \CS(G)$ is a linear character sheaf on $G$.
Then the restriction of $\cs{L}$ to $G\der$ is trivial if and only if $\cs{L} \cong q^*(\cs{L}\ab)$ for some
character sheaf $\cs{L}\ab$ on $G\ab$.
\end{proposition}

@@

\begin{proposition} \label{prop:Gdiagram}
A linear character sheaf $(\cs{L}, \mu)$ on a smooth group scheme $G$ is commutative if and only if
\begin{enumerate}
\item $\cs{L} \cong q^* \cs{L}\ab$ for some character sheaf $\cs{L}\ab$ on $G\ab$,
\item the following diagram of Weil sheaves on $G \times G$ commutes.
  \[
  \begin{tikzcd}[row sep=30]
   m^*\cs{L} \arrow{d}[swap]{(q\times q)^*\xi\ab} \arrow{r}{\mu} & \cs{L}\boxtimes\cs{L} \arrow{d}{\vartheta}\\
   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} &  \theta^*(\cs{L}\boxtimes\cs{L})
  \end{tikzcd}
  \]
\end{enumerate}
\end{proposition}

@@

Let $G$ be as above.
Let $H$ be a group scheme and write $n : H \times H \to H$ for the multiplication morphism.
Let $a : H \times G\to G$ be a group action compatible with the group structure on $G$ and write $p : H\times G \to G$ for projection. 
Consider the morphisms
\[
\begin{tikzcd}
H \times H \times G 
\arrow[shift left=4]{r}{b_1, b_2, b_3} 
\arrow{r}{}
\arrow[shift right=4]{r}{} 
& H \times G 
\arrow[shift left=4]{r}{a}
\arrow[shift right=4]{r}[swap]{p}
 & G 
\end{tikzcd}
\]
defined by $b_1(h_1,h_2,g) = (n(h_1,h_2),g)$,  $b_2(h_1,h_2,g) = (h_1,a(h_2,g))$ and $b_3(h_1,h_2,g) = (h_2,g)$.
Then, $a\circ b_1 = a\circ b_2$ and $a\circ b_3 = p\circ b_2$ and $p\circ b_1 = p\circ b_3$.
An $H$-equivariant Weil local sytem on $G$ is a Weil local system $\cs{L}$ on $G$ together with an isomorphism  
\[
\nu : a^*\cs{L} \to p^*\cs{L}
\] 
of Weil local systems on $H\times G$ such that the following diagram of isomorphisms of local systems on $H\times H\times G$ commutes,
\[
\begin{tikzcd}
\ &  \arrow{dl}[swap]{a\circ b_1 = a\circ b_2}  b_2^*  a^*\, \mathcal{L} \arrow{rr}{b_2^*(\nu)} && b_2^* p^*\, \mathcal{L} \arrow{dr}{p\circ b_2 = a\circ b_3} & \\
b_1^* a^*\, \mathcal{L} \arrow{dr}[swap]{b_1^*(\nu)} &&&&   \arrow{dl}{b_3^*(\nu)} b_3^* a^*\, \mathcal{L} \\
& b_1^* p^*\, \mathcal{L}  && \arrow{ll}[swap]{p\circ b_3 = p\circ b_1} b_3^* p^*\, \mathcal{L}, & 
\end{tikzcd}
\]
and $s^*(\nu) = \id_{\cs{L}}$.
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations above; so, for example, the top left isomorphism comes from $b_1^* a^* \iso b_2^* a^*$ since $a\circ b_1 = a\circ b_2$.
\iffalse
We remark that, on stalks, this condition is ...
\[
\begin{tikzcd}
\ &  \cs{L}_{x\cdot(y\cdot z)} \arrow{r}{\nu_{x,y\cdot z}} & \cs{L}_{y\cdot z} & \\
\cs{L}_{(x\cdot y)\cdot z} \arrow{dr}[swap]{\nu_{x\cdot y,z}} \arrow[equal]{ur} &&& \arrow[equal]{ul}  \arrow{dl}{\nu_{y,z}} \cs{L}_{y\cdot z} \\
& \cs{L}_z \arrow[equal]{r} & \cs{L}_z & 
\end{tikzcd}
\]
\fi
%
Morphisms of $H$-equivariant Weil local systems $(\cs{L},\nu)\to (\cs{L}',\nu')$ are morphisms of Weil local systems $\alpha: \cs{L}\to \cs{L}'$ for which the diagram
\[
\begin{tikzcd}
\arrow{d}[swap]{\nu} a^*\cs{L} \arrow{r}{a^*(\alpha)} & a^*\cs{L}' \arrow{d}{\nu'} \\
p^*\cs{L} \arrow{r}{p^*(\alpha)} & p^*\cs{L}'
\end{tikzcd}
\]
commutes.
This defines $\Loc_H(G)$, the category of $H$-equivariant Weil local systems on $G$.

@@

Now suppose $q : G \to G'$ is a regular epimorphism with kernel pair $(a,p)$ in the category of schemes
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=4]{r}{a}
  \arrow[shift right=4,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G'
\end{tikzcd}
\]
Let $\cs{L}'$ be a linear character sheaf on $G'$.
Consider the functor
\[
q^* : \CS(G') \to \CS(G)
\]
given on objects by $(\cs{L}',\mu') \mapsto (q^*\cs{L}', (q\times q)^* \mu', \nu_{\cs{L}'})$; cf [Dictionary].
Then $q^*\cs{L}'$ is an $H$-equivariant linear character sheaf on $G$ when equipped with the canonical isomorphism
\[
\nu_{\cs{L}'} : a^* q^* \cs{L}' \mathop{\longrightarrow}\limits^{\iso} (q\circ a)^* \cs{L}' = (q\circ p)^*\cs{L}' \mathop{\longrightarrow}\limits^{\iso} p^* q^*\cs{L}.
\]
This simple observation shows that the functor $q^* : \CS(G') \to \CS(G)$  factors according to the following commuting diagram of functors
\[
\begin{tikzcd}
\CS(G) &&\arrow{ll}[swap]{q^*} 
\CS(G') \arrow{dl}{\text{`comparison'}}\\
& \arrow{ul}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\]

@@

Suppose there is some $f : H\times G\to G\times G$ such that the following diagram commutes.
%%%%%%%% FIX!!!! %%%%%%%%%%
%\[
%\begin{tikzcd}
%&  \arrow{dl}[swap]{a} H \times G \arrow{dr}{p} \arrow{dd}[swap]{f} & \\
%G & & G \\
%& \arrow{ul}{m} G \times G \arrow{ur}[swap]{p} & 
%\end{tikzcd}
%\]
Then $f(h,g) = (f_1(h,g),p)$ for $f_1 : H\times G \to G$.
Let $\CS'(G)$ be the full subcategory of $\CS(G)$ consisting of those $(\cs{L},\mu,\nu) \in \CS_H(G)$ such that $f_1^*\cs{L} \iso (\EE)_{H\times G}$.

@@

\subsection{Proofs}

We can now give the missing proofs from Section \ref{sec:defs}.

\begin{proof}[Proof of Proposition~\ref{prop:Gder_triv}]
We now consider a special case of the lemmas above. 
Suppose $H = G\der$ and the action $a$ is the restriction of $m$ to $G\der\times G$.
Then $f : G\der \times G\to G\times G$ is simply the inclusion and $f_1: G\der\times G \to G$ is projection to the first coordinate followed by inclusion $G\der \hookrightarrow G$.

As we have seen above, the fppf quotient $q : G\to G\ab = G/G\der$ exists in schemes and is a $G\der$-torsor in the fppf topology. 
%(In fact, it's a trivial torsor in that topology.)
It follows from Lemma~[] that the comparison functor $\CS(G\ab)\to \CS_{G\der}(G)$ is an equivalence.
On the other hand, by Lemma~[], the forgetful functor $\CS_{G\der}(G)\to \CS(G)$ defines an equivalence $\CS_{G\der}(G)\to \CS'(G)$, where $\CS'(G)$ is the category of $(\cs{L},\mu,\nu)\in \CS_{G\der}(G)$ such that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$.
It now follows that $q^* : \CS(G\ab) \to \CS(G)$ is fll and faithful with essential image $\CS'(G)$.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:Gdiagram}]

\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:Gab}]
Combining Lemmas~[] and [] we see that $q^* : \CS(G\ab) \to \CS(G)$ is full and faithful.
% with essential image $\CS'(G)$, the category of linear character sheaves $\cs{L}$ on $G$ for which $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$. 
By Definition~[], this induces an equivalence between $\CCS(G\ab)$ and $\CCS(G)$. 
\end{proof}
